Utilizing Convolution for Obstacle Avoidance during Robotic Manipulator Movement

### 2023-07-21 Articles/ERK/erk.tex
ERK citations

In this research, we specifically consider a 7-DOF Panda Emika manipulator. When controlling both the position and orientation of the end-effector in a 3D space, which requires six degrees of freedom (three for position and three for orientation), a 7-DOF manipulator such as the Panda Emika has one redundant degree. However, when controlling only the position, which requires three degrees of freedom, the manipulator possesses four redundant degrees. This redundancy grants the manipulator an essential characteristic: an infinite number of possible joint configurations to reach a specific end-effector (EE) point in space.

@@

This paper proposes a novel method for real-time obstacle avoidance in the Panda Emika 7-DOF manipulator. We use an obstacle grid combined with convolution techniques to assign avoidance directions to each manipulator segment, ensuring efficient and safe maneuvering within complex environments.

@@

Two commonly employed methods for obstacle avoidance include potential field methods and artificial neural networks. In potential field methods, the robot and obstacles are treated as charged entities. The robot is attracted to the target (goal) while being repelled by obstacles, resulting in a potential field. The robot navigates this field, moving along the path of steepest descent. While this method is intuitive and relatively easy to implement, it can sometimes result in local minima problems where the robot gets stuck in a position that is not the target but cannot find a path due to the repelling obstacles.

@@

An advanced version of ANNs in the context of 3D obstacle detection and avoidance is the use of 3D Convolutional Neural Networks (3D-CNNs). While not strictly a path planning method, 3D-CNNs can be trained to recognize obstacles and free space in 3D occupancy data. The convolutional layers in the network can learn spatial hierarchies and structures, and the final layers can output a direction or action to take. These methods, although effective, require extensive training data and computational resources, however it is capable to recognizing and avoiding some local minima.

The Vector Field Histogram (VFH) method presents another approach. This method uses a 2D histogram grid as a spatial representation of the environment. The histogram cells are updated based on the proximity and angle of obstacles. Then a primary polar histogram is calculated from the grid cells, leading to a candidate valley for navigation. While originally developed for 2D, this method was later extended to 3D as VFH+.

Our proposed approach uses a 3D occupancy or voxel grid and 3D directional kernels, akin to Sobel filters, to perform kernel convolution and calculate the direction in which to move the robot arm to avoid obstacles. This method marries the benefits of convolutional approaches with the intuitive spatial reasoning provided by a voxel grid representation.

@@

\bibitem{Nguyen1991} Nguyen, Charles C., Zhou, Zhen-Lei, Mosier, Gary E. (1991) Analysis and control of a kinematically redundant manipulator. \textit{Computers \& Electrical Engineering}, 17(3), pp. 147–161. doi:10.1016/0045-7906(91)90031-T.
\bibitem{ERK} ERK, http://www.ieee.si/erk/index.html 
\bibitem{Zbornik} B. Zajc, A. Trost: Zbornik triindvajsete mednarodne Elektrotehniške in računalniške konference ERK 2014, 22. - 24. September 2014, Portorož, Slovenija

>>>>

In this research, we specifically consider a 7-DOF Franka Panda Emika~\cite{Franka2021} manipulator. When controlling both the position and orientation of the end-effector in a 3D space, which requires six degrees of freedom (three for position and three for orientation), a 7-DOF manipulator such as the Panda Emika has one redundant degree. However, when controlling only the position, which requires three degrees of freedom, the manipulator possesses four redundant degrees. This redundancy grants the manipulator an essential characteristic: an infinite number of possible joint configurations to reach a specific end-effector (EE) point in space.

@@

This paper proposes a novel method for real-time obstacle avoidance of the manipulator arm. We use an obstacle grid combined with convolution techniques to assign avoidance directions to each manipulator segment, ensuring efficient and safe maneuvering within complex environments.

@@

Two commonly employed methods for obstacle avoidance include artificial potential field methods (APF)~\cite{Khatib1985} and artificial neural networks. In potential field methods, the robot and obstacles are treated as charged entities. The robot is attracted to the target (goal) while being repelled by obstacles, resulting in a potential field. The robot navigates this field, moving along the path of steepest descent. While this method is intuitive and relatively easy to implement, it can sometimes result in local minima problems where the robot gets stuck in a position that is not the target but cannot find a path due to the repelling obstacles.

@@

An advanced version of ANNs in the context of 3D obstacle detection and avoidance is the use of 3D Convolutional Neural Networks (3D-CNNs)~\cite{Duguleana2011}. While not strictly a path planning method, 3D-CNNs can be trained to recognize obstacles and free space in 3D occupancy data. The convolutional layers in the network can learn spatial hierarchies and structures, and the final layers can output a direction or action to take. These methods, although effective, require extensive training data and computational resources, however it is capable to recognizing and avoiding some local minima.

The Vector Field Histogram~\cite{Borenstein1991} (VFH) method presents another approach. This method uses a 2D histogram grid as a spatial representation of the environment. The histogram cells are updated based on the proximity and angle of obstacles. Then a primary polar histogram is calculated from the grid cells, leading to a candidate valley for navigation. While originally developed for 2D, this method was later extended to 3D as VFH+~\cite{Ulrich1998}.

Our proposed approach uses a 3D occupancy or voxel grid and 3D directional kernels, akin to 2D Sobel filters, to perform kernel convolution and calculate the direction in which to move the robot arm to avoid obstacles. This method marries the benefits of convolutional approaches with the intuitive spatial reasoning provided by a voxel grid representation.

@@

\bibitem{Nguyen1991} Nguyen, Charles C., Zhou, Zhen-Lei, Mosier, Gary E. (1991) Analysis and control of a kinematically redundant manipulator. \textit{Computers \& Electrical Engineering}, 17(3), pp. 147–161. doi:10.1016/0045-7906(91)90031-T.

\bibitem{Franka2021} Franka Emika. (2021). Panda: Specification. \textit{Available at: https://www.franka.de/panda}

\bibitem{Khatib1985} Khatib, O. (1985) Real-time obstacle avoidance for manipulators and mobile robots. In: \textit{1985 IEEE International Conference on Robotics and Automation Proceedings}, 2, pp. 500–505. doi:10.1109/ROBOT.1985.1087247.

\bibitem{Duguleana2011} 
M. Duguleana, F. G. Barbuceanu, A. Teirelbar, and G. Mogan,
"Obstacle avoidance of redundant manipulators using neural networks based reinforcement learning,"
\textit{Robotics and Computer-Integrated Manufacturing}, 
DOI: 10.1016/j.rcim.2011.07.004, Sep. 2011.

\bibitem{Borenstein1991} 
J. Borenstein and Y. Koren,
"The vector field histogram-fast obstacle avoidance for mobile robots,"
\textit{IEEE Transactions on Robotics and Automation}, 
vol. 7, no. 3, pp. 278–288, Jun. 1991.
DOI: 10.1109/70.88137

\bibitem{Ulrich1998} 
I. Ulrich and J. Borenstein,
"VFH+: reliable obstacle avoidance for fast mobile robots,"
In \textit{Proceedings. 1998 IEEE International Conference on Robotics and Automation}, 
vol. 2, pp. 1572–1577, May 1998.
DOI: 10.1109/ROBOT.1998.677362
